% acknowl.tex {Acknowledgements}

\addcontentsline{toc}{chapter}{Acknowledgements}

\chapter*{Acknowledgements}

First and foremost, I would like to thank God for making this possible, without him i will not be where i am today. I would also like to thank my parents, Benneth and Chinwe Nwafor for their constant encouragement, love and support. I would like to thank my advisor Dr. Gedare Bloom for his guidance and encouragement throughout this process, for believing in my research ideas and also for his constructive criticism and valuable input in various drafts revisions of my dissertation. I would like to thank my committee members Dr. Rawat Danada, Dr. Legand Burge, and Dr. Charles Kim for their insightful comments and feedback. I would like to thank my God mother Dr.(Mrs) Ifeoma Obiwuru for her unwavering love and support throughout my academic career and for also encouraging me to reach greater heights. I would like to thank some of the most amazing mentors i have been fortunate to have: Dr. Adedoyin Adeyoiga, and Dr. Deivy Peterescu. You all are a source of inspiration and a beacon of light. You have provided me with valuable advice throughout my academic career. Finally, I would like to thank everyone whose name i might have not mentioned that have played a part in the successful completion of my doctoral dissertation.



This material is based upon work supported by the National Science
Foundation under Grant No. CNS-1646317 and the U.S. Department of
Homeland Security under Grant Award Number 2017-ST-062-000003. Any
opinions, findings, and conclusions or recommendations expressed in
this material are those of the authors and do not necessarily reflect
the views of the National Science Foundation and should not be
interpreted as necessarily representing the official policies, either
expressed or implied, of the U.S. Department of Homeland Security.